An alternative MIP formulation for the Military Flight and Maintenance Planning problem

### 2020-02-28 R/presentations/roadef202002/presentation.Rmd

@@

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenances operations** have specific rules that govern their frequency and capacity.

@@

.center[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
library(RefManageR)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
knitr::include_graphics('calendar3.png')

@@

* A series of $j \in \mathcal{J}$ missions are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all missions are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each mission requires a certain number $r_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $h_j$ hours in each period. An aircraft assigned to a mission $j$ must be assigned for at least $MT^{min}_{j}$ and at most $MT^{max}_{j}$ consecutive periods.
* The heterogeneous fleet requires recurrent preventive maintenance operations (checks) as a result of the total flight hours since the last check. There are also min (max) distances between maintenances.
* A check takes exactly $m$ periods and cannot be interrupted. There is a limited capacity for maintenances at each period.
* The objective is to minimize the total number of checks while maximizing the status of the fleet at the end of the horizon.

@@

* $a_{jti}$   :  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
* $a^s_{jti}$ :  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$.  
* $m_{it}$    :  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.  
* $u_{it}$    :  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$.

\begin{align}
 & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
 & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
 & rft_{it} \leq rft_{i(t-1)} + H^M m_{it} - u_{it}
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:first_rft_upper}\\
& rft_{it} \in [0,H^M]
      & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:first-mu}
\end{align}

@@

\begin{align}
& \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
  & \hspace{200px}  i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \label{eq:cycle_hours2}\\
\end{align}

@@

\begin{align}
  & \text{Max}\;
  \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
  & \sum_{i \in \mathcal{I}, p \in \mathcal{P}_{t}} m_{ip} \leq C^{max} 
          & t \in \mathcal{T} \label{eq:capacity1}\\
  & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
          & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
  & \sum_{p \in \mathcal{P}_{t}} m_{ip} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
          & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
    & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
\end{align}

@@

.pull-left[
* The distance between maintenance has a maximum of $E^{max}$ periods.
* Depending on the instance, the optimal distance can be shorter.
* This distance conditions the total number of patterns to create.
]

@@

### Forecasting + Optim

@@

1. Create forecasting model based in 5000 small instances.
2. Use forecasting model to predict bounds on distance between maintenaances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
3. Implement the pseudo-cut:

@@

@@

* **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.
* **Warm-start Column Generation** with a selected subset of potentially good patterns.
* **Automatize prediction** so it can be easily integrated in other problems.

>>>>

```{r, echo=FALSE, warning=FALSE}
library(RefManageR)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

```

@@

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenance operations** have specific rules that govern their frequency and capacity.

@@

.center[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('calendar3.png')

@@

* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the end of the horizon.

@@

* $a_{jti}$   :  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
* $m_{it}$    :  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.  
* $u_{it}$    :  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$.

\begin{align}
 & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
 & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
 & rft_{it} \leq rft_{i(t-1)} - u_{it} + H^M m_{it}
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:first_rft_upper}\\
& rft_{it} \in [0,H^M]
      & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:first-mu}
\end{align}

@@

\begin{align}
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{M} + M (1 - m_{ip}) & \notag \\
  & \hspace{200px}  i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \label{eq:cycle_hours2}\\
\end{align}

```{r, echo=FALSE, out.width='100%', warning=FALSE}
knitr::include_graphics('calendar3.png')

```

@@

\begin{align}
  & \text{Max}\;
  \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
  & \sum_{i \in \mathcal{I}, p \in \mathcal{P}_{t}} m_{ip} \leq C^{max} 
          & t \in \mathcal{T} \label{eq:capacity1}\\
  & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
          & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
  & \sum_{p \in \mathcal{P}_{t}} m_{ip} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
          & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^M + M (1 - m_{ip}) & \notag \\
    & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
\end{align}

### Is this better?

1. It uses 3 times the number of constraints and 3 times the number of variables.
  * variables: 11000 => 28000.
  * constraints: 13000 => 48000.
1. It is still better. Better lineal relaxation, better performance.

@@

.pull-left[
* The distance between maintenance has a maximum of $E^{M}$ periods.
* Depending on the instance, the optimal distance can be shorter.
* This distance conditions the total number of patterns to create.
]

@@

### Forecasting + Optimization

@@

1. Create forecasting model based in 5000 small instances.
2. Use forecasting model to predict bounds on distance between maintenances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
3. Implement the pseudo-cut:

@@

4. Recycling.

@@

* **Better predictions** with better features, or predicting several characteristics of optimal solutions.
* **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.
* **Warm-start Column Generation** with a selected subset of potentially good patterns.
* **Automatize prediction** so it can be easily integrated in other problems.
